---
title: "Label line ends in time series"
author: "Valerie Egoshin"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    css: spfiles/style.css
    toc: yes
    toc_float: yes
    fig_caption: yes
    fig_width: 6
    fig_height: 4
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

 
#### Label line ends in time series with ggplot2


From [r-blogger](https://www.r-bloggers.com/label-line-ends-in-time-series-with-ggplot2/) or [original](https://drsimonj.svbtle.com/label-line-ends-in-time-series-with-ggplot2)


### Data

The Orange data frame has 35 rows and 3 columns of records of the growth of orange trees. 

* Tree
    * an ordered factor indicating the tree on which the measurement is made. The ordering is according to increasing maximum diameter.
* age
    * a numeric vector giving the age of the tree (days since 1968/12/31)
* circumference
    * a numeric vector of trunk circumferences (mm). This is probably “circumference at breast height”, a standard measurement in forestry.


```{r data_}
d <- Orange
sample_n(d, 6)

count(d, Tree)

```


### Template code
To create the basic case where the numbers appear at the end of your time series lines, your code might look something like this:


```{r template, eval=FALSE}
# You have a data set with:
# - GROUP colum
# - X column (say time)
# - Y column (the values of interest)
DATA_SET

# Create a vector of the last (furthest right) y-axis values for each group
DATA_SET_ENDS <- DATA_SET %>% 
  group_by(GROUP) %>% 
  top_n(1, X) %>% 
  pull(Y)

# Create plot with `sec.axis`
ggplot(DATA_SET, aes(X, Y, color = GROUP)) +
    geom_line() +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(sec.axis = sec_axis(~ ., breaks = DATA_SET_ENDS))

```

### Let's see it

Let’s break it down a bit. We already have our data set where the group colum is Tree, the X value is age, and the Y value is circumference.

`dplyr::pull` - The default returns the last column (on the assumption that's the column you've created most recently)

```{r template_result}
# So first get a vector of the last (furthest right) values for each group:
d_ends <- d %>% 
  group_by(Tree) %>% 
  top_n(1, age) %>% 
  pull(circumference) 

d_ends 

```

### Basic plot

```{r fig1, fig.cap='Basic view'}
ggplot(d, aes(age, circumference, color = Tree)) +
      geom_line()

```

<br>
Now we can use `scale_y_*`, with the argument `sec.axis` to create a `second axis` on the right, with numbers to be displayed at breaks, defined by our vector of line ends:

```{r fig2, fig.cap='Basic plot plus second axis on the right'}
ggplot(d, aes(age, circumference, color = Tree)) +
      geom_line() +
      scale_y_continuous(sec.axis = sec_axis(~ ., breaks = d_ends))

```

The only major addition I suggest is expanding the margins of the `x-axis` so the gap disappears. You do this with `scale_x_*` and the `expand` argument:

```{r fig3, fig.cap='Some improvements'}
ggplot(d, aes(age, circumference, color = Tree)) +
      geom_line() +
      scale_y_continuous(sec.axis = sec_axis(~ ., breaks = d_ends)) +
      scale_x_continuous(expand = c(0, 0))

```

### Polishing it up

```{r fig4, fig.cap="Final plot"}
d %>% 
  ggplot(aes(age, circumference, color = Tree)) +
    geom_line(size = 2, alpha = .8) +
    theme_minimal() +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(sec.axis = sec_axis(~ ., breaks = d_ends)) +
    ggtitle("Orange trees getting bigger with age",
            subtitle = "Based on the Orange data set in R") +
    labs(x = "Days old", y = "Circumference (mm)", caption = "Plot by @drsimonj")

```

